Due to its novelty, it is necessary to demonstrate that \ac{osier} generates
reproducible and reliable answers consistent with the results from an
established framework and that \ac{osier} is internally consistent when applying
different methods. To that end, this chapter illustrates some of the differences
between the logical and optimal dispatch algorithms introduced in Section
\ref{section:dispatch_model} by testing the two dispatch algorithms  
on some simple cases, as well as comparing the time scaling of these two
methods. Second, this chapter verifies that solutions calculated by \ac{osier}
agree with a more established \ac{esom}, \ac{temoa} \cite{hunter_modeling_2013}.
Finally, this chapter shows some of \ac{osier}'s advanced features, such as
many-objective objective problems and combining \ac{moo} with \ac{mga}. The next
section introduces some of the data and methods used to prepare the later
example problems.


\section{Technology Data}
\input{4-benchmark-results/40-1-technology-data}
\section{Data for the dispatch algorithm comparison}
The following sections describe the synthetic data generated to run the experiments.

\input{4-benchmark-results/41-1-example-data}
\section{Comparing dispatch algorithms}
The first exercise in this section directly compares the results of two dispatch
algorithms introduced in Section \ref{section:dispatch_model} to confirm that
the two algorithms give similar answers. The next exercise shows how the two
algorithms scale. The final exercise focuses on the logical dispatch algorithm
and investigates how that algorithm scales with the number of threads used for
parallelization.
% \input{4-benchmark-results/41-demonstrate}
\input{4-benchmark-results/45-new-dispatch}
\section{Benchmarking \ac{osier}}
The previous sections demonstrated internal consistency between two dispatch algorithms
within \ac{osier} and showed how these algorithms scale with problem size and the number
of available threads. This section explores the performance different \acp{ga}, compares 
\ac{osier} to the established \ac{esom}, \ac{temoa}, and shows how \ac{osier} exceeds
current state-of-the-art.
\input{4-benchmark-results/42-1-benchmark-data}
\input{4-benchmark-results/42-decide-ea}
\input{4-benchmark-results/43-explore}
\input{4-benchmark-results/44-four-obj}
