Every year, world leaders meet to discuss plans to address climate change at the
COP summit. In 1995, world leaders established a set of targets with the Kyoto
Protocol \cite{united_nations_kyoto_1998} and again with the 2016 Paris Climate
Agreement \cite{united_nations_paris_2015}. Every few years, the United Nations
releases a report from the \ac{ipcc} assessing the current impacts of climate
change and forecasting future scenarios. Most of the world understands that
anthropogenic climate change is an existential threat to society. Indeed, many
studies in the \ac{esom} literature begin with a statement about the urgency of
climate change. This chapter reviews the extant literature for both quantitative
and qualitative analyses of the problem considered in this thesis --- primarily
bridging the gap between feasibility or planning studies to address the climate
crisis and the current pattern of missed targets and growing carbon emissions.
First, I draw from the risk assessment literature to characterize and situate
the problem of climate change and demonstrate the necessity of a holistic
analysis. Second, I build upon the central issue of disproportionality of
climate change risk by reviewing the energy and environmental justice
literature. Third, I develop an encompassing definition of an ``energy system''
using technical and social perspectives. Finally, I review the energy system
literature for gaps in conventional modeling practices and identify previous
attempts to incorporate social science and justice concepts into energy system
models.

% The first section reviews methods and attempts to model energy systems for
% planning purposes. The second section reviews the ways social movements help
% or hinder various energy projects, how governments succeed or fail to achieve
% their energy goals from this perspective, and why some communities favor or
% disfavor energy projects.

\input{2-literature/23-social-movements}
% \input{2-literature/26-nuclear-energy} 
% \input{2-literature/27-renewables}
% \input{2-literature/28-procedures}
\input{2-literature/25-energy-justice}
\input{2-literature/21-esoms}
\input{2-literature/22-moop}
\textcolor{red}{\section{Hyper-local energy system modeling}}
\textcolor{red}{To the literature review, I want to add a section discussing 
the research on modeling at the municipal level. These papers start with the 
following: 
\cite{mckenna_combining_2018,johannsen_municipal_2023,ben_amer_too_2020}.}
\input{2-literature/24-justice-esoms}
