\section{Discussion}
This section synthesizes the interview results into recommendations that can be
applied in other contexts.

\subsection{Who can benefit from energy modeling?}
Much of the literature that exists on modeling practices at the municipal scale
find that \acp{esom} are seldom used
\cite{ben_amer_too_2020,johannsen_municipal_2023}. The reasons cited for this
are \cite{ben_amer_too_2020,johannsen_municipal_2023,susser_better_2022}
\begin{enumerate}
    \item poorly scoped models that do not capture relevant phenomena,
    \item lack of resources or expertise to conduct modeling, 
    \item and lack of collaboration between modelers and planners.
\end{enumerate} 
Specifying challenges in this way provides a list of issues that can be
addressed. Unfortunately, the interviews in this study suggest that even if each
of these issues were appropriately handled that municipalities would still not
benefit from the use of energy models. Instead, there are structural barriers to
the municipal use of energy models. The typical focus of \acp{esom} on national
and regional energy systems evinces an implicit understanding of these barriers.
Municipalities can still benefit from mathematical optimization tools as long as
a municipality has sufficient autonomy over its energy choices to impact the
decision variables in the model. For example, a municipality that does not own
its electric utility nor any generating resources would not benefit from a
capacity expansion model that optimizes investments in generating resources. Of
the municipalities considered in this study, \ac{uiuc} satisfies this condition
the best and is the most likely to benefit from \acp{esom} in planning
processes. Indeed, meeting its climate goals might necessitate modeling.

Although municipal governments may not be the strongest candidates for energy
modeling, nonprofits and grassroots organizations could benefit from energy
modeling to aid the development of their energy visions which can then be used
to lobby for policy changes. For example, with some support \ac{nest} could use
a model to create an alternative energy vision than the one provided by
\ac{imea}.

\subsection{How \ac{osier} addresses the ``human dimension''} 

\ac{osier} allows for any number of constraints, objectives, and even allows for
user-defined objectives, a design philosophy consistent with the first principle
from Elmallah et al. which called for ``place-based'' solutions
\cite{elmallah_frontlining_2022}. However, place-based solutions require
authentic engagement and ``quantitative metrics cannot replace on the ground
work that includes qualitative, community-based approaches in devising policy
solutions'' \cite{elmallah_frontlining_2022}. While \ac{osier} can stand alone
as an \ac{esom}, it is likely most effective as part of an iterative,
participatory, approach as outlined by McGookin et al. 2024
\cite{mcgookin_advancing_2024}. This approach breaks down the modeling process
into five distinct parts
\begin{enumerate}
    \item Research design,
    \item Model assumptions,
    \item Modelling results,
    \item Outreach and communication,
    \item and Evaluation.
\end{enumerate}
Each part in this process can be iterated upon based on the outcomes from a
later step \cite{mcgookin_advancing_2024}. \ac{osier} impacts the model
assumptions, and modeling results stages the most since model objectives become
assumptions when using \ac{osier} and understanding the tradeoffs of different
model solutions becomes part of the discussion. Analyzing tradeoffs is important
for reducing the normative uncertainty present in all modeling exercises and
policy discussions. The interviews in this study support \ac{osier}'s ability to
facilitate dialogue and communicate with decision makers and constituents.
Tradeoff analysis was identified as a particularly important feature of the
tool.

\subsection{Policy Recommendations}
Although the state of Illinois uses modeling results and hires consultants to
perform modeling, the responses from interviewees suggest that even more
transparency in the model data and the models themselves are desired. A public
comment period does not constitute genuine involvement of a participatory
process. Creating such a process would align Illinois more closely with the
energy justice planning principles outlined by Elmallah et al.
\cite{elmallah_frontlining_2022}. Specifically ``shifting the balance of power
from existing forms of energy governance.'' Illinois has an opportunity to
employ participatory modeling to develop robust visions for its energy future.
To that end, I suggest Illinois create such a process and use open source
modeling tools whenever possible.

\subsection{Model and Data Development}
Several of the model features requested by the interviewees would be better
suited by a different modeling paradigm than the one currently provided by
\ac{osier}. For example, questions about technology adoption rates, testing
incentive structures, or time of use rates --- issues related to human behavior
--- are likely better addressed by \acp{abm}. Questions about energy efficiency
might be better served by 1) \acp{bem} but also 2) better data about the cost of
efficiency improvements. \ac{nrel} has a large amount of simulated building
level energy use data along with the performance of various efficiency upgrades
but no information on cost \cite{wilson_end-use_2022}. \ac{osier} could readily
represent efficiency as a ``technology'' option, but this would require capital
cost data on a per kW basis. Some of this data exists but
the data have large uncertainties \cite{less_cost_2021}.


\subsection{Limitations of this study}
This study has a couple of limitations. First, despite the varied contexts all
of the study participants have an Illinois perspective. Illinois is a
progressive state with a deregulated electricity market. Municipalities in less
progressive states or states with vertically integrated electricity markets may
have even less choice over their energy supply than in Illinois. Second,
\ac{osier} was already mostly complete at the time of this study. Ideally, this
study could have been a co-creative process that developed a tool with the
features requested by city planners. Interviewees also had limited exposure to
\ac{osier} before offering their insights. Lastly, this study did not actually
employ \ac{osier} in a planning or energy visioning process.