Writing a thesis can often feel like a solitary experience. Despite appearances,
such an endeavor is only possible through the collective support of mentors,
colleagues, friends, and family (and those friends who became family).

I am profoundly indebted to my advisor and mentor, Professor Kathryn D. Huff,
who set me on this path. Her uncompromising standards for excellence are matched
only by the care and attention she affords her students. I only hope I can live
up to her example. I would also like to thank my co-advisor, Professor Madicken
Munk, for encouraging me to pursue my curiosity, which led to this sprawling
multi-disciplinary work.

I thank the other members of my committee, James Stubbins, Clifford Singer,
McKenzie Johnson, and Denia Djoki\'c. I am particularly grateful to Professor
Johnson for her patience and humor while mentoring me through the nuances of
interacting with and writing about real people (gasp!). I am also deeply
appreciative of the time and effort Dr. Djoki\'c put into guiding myself and
Haley Williams (UC Berkeley) through the philosophy of science literature in
weekly meetings of the ``Nuclear Contextualizers."

Of course, I thank my colleagues, past and present, from the Advanced Reactors
and Fuel Cycles (ARFC) group: Gwendolyn Chee, Sun Myung Park, Amanda Bachmann,
Nathan Ryan, Olek Yardas, Zo\"e Richter, and Luke Seifert. Their feedback and
code reviews were indispensable. 

I would like to extend a special thanks to the following people. Shannon
Anderson, thank you for helping me navigate the challenges of social science and
engaging with the energy justice literature. Roberto Fairhurst, thanks for being
a great friend and roommate. I miss interrupting you at work (and I appreciate
your patience with my interruptions). Jeremy Mettler, although I credit
Professor Huff with officially starting my journey, I truly would not be here
without your persistent, friendly nudges. Thank you for always inviting me, for
being a trusted confidant, and a dear friend. Nataly Panczyk, I adore your
curiosity, kindness, and infectious enthusiasm. I am perpetually in awe of you. 

I would also like to acknowledge the many incredible professors at the
University of Illinois and at the College of Lake County. In particular, I
acknowledge Drs. Ryan Stock, Ana Mazilu, Josephine Faulk, Kristi Dameron, and
Steve Chamberlin for encouraging me to pursue graduate studies.

Finally, I received financial support for this work from the Nuclear Regulatory
Commission Fellowship Program, the Felix T. Adler Fellowship from the NPRE
Department at UIUC, and the Union of Concerned Scientists Professional
Development Fund.